Learning for Logistic Regression Model in Parallel

### 2013-06-22 draft.tex
IEEE version revised

% This is paper draft
% edited by Haoruo Peng
% Date: 2013.6.14
\documentclass[10pt, conference, compsocconf]{IEEEtran}
\usepackage{amsmath, amssymb}
\usepackage{color}
\usepackage{graphicx}
\usepackage{bbding}
\usepackage{epstopdf}

@@

%\newtheorem{theorem}{Theorem}
%\newtheorem{lemma}{Lemma}
%\newtheorem{definition}{Definition}
%\newtheorem{proposition}{Proposition}
%\newtheorem{corollary}{Corollary}
%\newtheorem{example}{Example}

\begin{document}
%\mainmatter  % start of an individual contribution

% first the title is needed
\title{Learning for Penalized Logistic Regression Model in Parallel}

% a short form should be given in case it is too long for the running head
%\titlerunning{Learning for Penalized Logistic Regression Model in Parallel}

\author{\IEEEauthorblockN{Authors Name/s per 1st Affiliation (Author)}
\IEEEauthorblockA{line 1 (of Affiliation): dept. name of organization\\
line 2: name of organization, acronyms acceptable\\
line 3: City, Country\\
line 4: Email: name@xyz.com}
\and
\IEEEauthorblockN{Authors Name/s per 2nd Affiliation (Author)}
\IEEEauthorblockA{line 1 (of Affiliation): dept. name of organization\\
line 2: name of organization, acronyms acceptable\\
line 3: City, Country\\
line 4: Email: name@xyz.com}
}
%
%\author{Haoruo Peng\inst{1, 2}  \and Ding Liang\inst{1} \and Cyrus Choi\inst{1} \and Edward Y. Chang\inst{1} \and Yin Zhao\inst{2}}
%
%\authorrunning{Haoruo Peng \and Ding Liang \and Cyrus Choi \and Edward Y. Chang \and Yin Zhao}
% (feature abused for this document to repeat the title also on left hand pages)

%\institute %
%{
%\inst{}
%HTC Research, Beijing, China 100084 \\
%\and
%\inst{}
%Department of Computer Science and Technology\\
%Tsinghua University, Beijing, China 100084
%\email{penghaoruo@hotmail.com ...}
%}
%
%\toctitle{Lecture Notes in Computer Science}
%\tocauthor{Authors' Instructions}
\maketitle

\begin{abstract}
Logistic regression (LR) model has a variety of applications in the field of machine learning.
It is a supervised learning model, and can be extended to models with regularization, which enjoy the benefits of solid statistical base.
This paper addresses the issue of computational efficiency for solving LR in big data scenario.
We focus on both algorithm level and computation platform level of task parallelism.
We study three different parallel computation platforms along with three different types of parallel algorithms to efficiently improve scalability.
Hadoop, Mahout and Spark are three existing well-known parallel computing platforms that can be implemented with large scale machine learning algorithms.
Parallel gradient Descent and stochastic gradient descent are two state-of-art algorithms for parameter optimization in LR.
We also present a novel parallel sublinear method based on its sequential version.
We then make a comparison between these algorithms implemented in different platforms.
As the outcome shows, we can choose different algorithms in different situations in terms of datasets and machine resources for tradeoff between efficiency and precision.
To enhance practicality, we provide fault tolerance for those  parallel distributed computations.
\end{abstract}

@@

\section{Introduction} \label{sec:int}
Logistic regression model~\cite{HastieBook:SL} plays a vital role in machine learning and data mining.
It is a widely used model in algorithms like PageRank~\cite{page1999pagerank} and Anti-spam Filtering~\cite{androutsopoulos2000evaluation}.
The model serves for classification problems, and enjoys a substantial body of supporting theories and algorithms.
Thus, a binary classification problem modeled by LR can be easily extended to a multi-classification problem.
Also, due to the strong statistical foundation of LR model, we can extend results into a deeper analysis which may apply to other classification models.
We will focus on the binary LR model in the rest of the paper.

In recent years, many modern massive datasets are getting larger and larger.
A most identifiable characteristic of these datasets is that the size of their training data is very large and data dimensionality is very high.
Large data volume and high data dimensionality bring unavoidable computational challenges to machine learning problems.
For example, in social networks, we have datasets~\cite{lewis2008tastes} covering multiple dimensions like friendship relationship, share of common interests. They generally include of millions of records over millions of attributes. More evidently, for the task of text categorization, we usually have to face a dataset with data volume in the billion scale and each data instance is characterized by thousands of feature dimensions~\cite{lewis2004rcv1}.
We tackle these challenges on both algorithm level and computing platform level via parallelization methods.
For many applications, in order to have speedup in performance, we can employ not only task parallelism, but also data parallelism~\cite{subhlok1993exploiting}.
In this paper, we mainly discuss issues in task parallelism.

When developing parallel algorithms for LR, it is unavoidable to come across the question of choosing which parallel computing platform to employ.
After a thorough survey, we choose three unique and best-known computing platforms to test: Hadoop~\cite{white2012hadoop}, Mahout~\cite{mahoutscalable} and Spark~\cite{zaharia2010spark}.
Hadoop employs HDFS~\cite{borthakur2008hdfs} and MapReduce~\cite{dean2008mapreduce}.
Mahout runs on Apache Hadoop and makes improvements in the low level code.
Spark features for iterative algorithms and also supports HDFS.
We will compare design features of them separately in related work section.

In sequential parameter optimization algorithms for LR model, a classical way is to turn to stochastic approximation methods.
Stochastic approximation methods, such as stochastic gradient descent~\cite{zhang2004solving} and stochastic dual averaging~\cite{xiao2010dual}, obtain optimal generalization guarantees with a small number of passes over the data with runtime linear to the dataset size.
The stochastic gradient descent method can work in an online fashion, as we view the sampled data points as the come-in data stream.
It is hard to be changed to task-parallelism, but it is blazingly fast.
If we take a step back and go back to the general gradient descent to solve LR model, we can find that it can be actually embarrassingly parallel.
We can take in all the data at the same iteration and just compute the gradient in a MapReduce fashion.
We can further speed up the runtime by employing sublinear algorithms~\cite{peng2012sublinear} via the use of stochastic approximation idea.
This methods access a single feature of training vectors instead of entire training vectors at each iteration.
We propose a parallel version of this algorithm and achieve the result of comparable accuracy while being faster convergent.

The rest of the paper is organized as follows:
Section~\ref{sec:rew} discusses some related work.
In Section~\ref{sec:plr}, we review the logistic regression model and the sequential sublinear algorithm.
In Section~\ref{sec:framework}, we present the parallel framework of all three different algorithms for LR model.
Section~\ref{sec:setup} describes the datasets and the baseline of our experiments and Section~\ref{sec:experiment} presents the experimental results.
Finally, we offer our concluding remarks in Section~\ref{sec:concl}.

@@

\subsection{Scalability Improvement in Big Data Research}
In big data scenario, large datasets require us to develop machine learning algorithms towards a more efficient and more parallelized end.
Researchers have already done much work to cater for the needs of massive datasets.
Early work like PSVM (\textit{Parallel Support Vector Machines})~\cite{chang2011psvm} employed an approximation matrix decomposition method based on row selections to reduce the memory when running the algorithm.
It can then increase the number of participated parallel computation nodes to several hundreds.
Later, the work of PLDA (\textit{Parallel Latent Dirichlet Allocation})~\cite{wang2009plda} further improve the computational efficiency of the LDA model through the use of sampling methods.
The proposed parallel algorithm in Hadoop is robust, for it has fault tolerance of machine failures, taking the advantage of Hadoop.
Recently, the work by Dean et al.\cite{le2011building} showed that the advantage of parallelization is fully taken in deep learning algorithms.
It pushed the limits of parallel computational nodes to a hundred-million level, and in the meantime, achieved a best learning performance ever.
GraphLab~\cite{kyrola2012graphchi}, a recent developed tool for large-scale machine learning. It tackles with efficiency and scalability problems of machine leaning algorithms when applied to graphs.

\subsection{Three Computing Platforms}
Three computing platforms we are working on each features for its own advantage in implementing machine learning algorithms.
The Apache Hadoop~\cite{white2012hadoop} software library allows for distributed processing of large datasets across clusters of computers using simple programming models.
Hadoop utilizes MapReduce~\cite{dean2008mapreduce} as its computational paradigm.
MapReduce paradigm takes a set of key-value pairs as its input.
Then a user-defined \textit{map} function turns input key-value pairs into some intermediate key-value pairs.
A user-defined \textit{reduce} function merges values for the same intermediate key to produce output.
All of the processing is independent and it makes MapReduce paradigm easy to be parallelized.
In addition, Hadoop provides a distributed file system (HDFS).
Both MapReduce and HDFS are designed to handle node failures in an automatical way, making Hadoop support large clusters built on commodity hardware.
The goal of Apache Mahout~\cite{mahoutscalable} is to build scalable machine learning libraries. Their core algorithms for clustering, classification and batch based collaborative filtering are implemented on top of Apache Hadoop using the MapReduce paradigm. The core libraries are so highly optimized that it also has good performance for non-distributed algorithms.
Spark~\cite{zaharia2010spark} is a cluster computing system that aims to make data analytics fast.
Spark provides support in-memory cluster computing. A job can load data into memory and query it repeatedly by creating and caching resilient distributed datasets (\textit{RDDs}).
Moreover, RDDs achieve fault tolerance through lineage: RDDs always have enough information about how they were derived from other RDDs, and they are easy to be rebuilt if a certain partition is lost.
Spark can run up to 100 times faster than Hadoop MapReduce for iterative algorithms.

In Table~\ref{tab:systemcmp}, we compare between Hadoop and Spark computing platforms.
As Mahout is mainly built on Hadoop, it inherits most of Hadoop features.
To better illustrate, we use short abbreviations in the table.
Spark supports two types of operations on RDDs: Actions and Transformations (As and Ts).
Actions include functions like \textbf{count}, \textbf{collect} and \textbf{save}. They usually return a result from input RDDs.
Transformations include functions like \textbf{map}, \textbf{filter} and \textbf{join}. They normally build new RDDs from other RDDs, which are under the lineage rule.
In the table, we use "NFS" to represent "Normal File System" and "Fault Tolerance" here is specifically for node failure situations.
\begin{table}[h]
\centering
\caption{Platform Comparison: Hadoop VS Spark}\label{tab:systemcmp}
\begin{tabular}{|c|c|c|}
\hline
           & Hadoop & Spark \\
\hline
Computational Paradigm  & MapReduce & As and Ts \\
\hline
File System Supported & HDFS & HDFS and NFS \\
\hline
Design Concept  & Key-value Pairs & RDD \\
\hline
Fault Tolerance Technique & Redundancy & Lineage \\
\hline
\end{tabular}
\end{table}

\subsection{Previous Research on Sublinear Method}
Sublinear method is recently developed by many contributing researchers.
Clarkson et al.~\cite{clarkson2010sublinear} first proposed the method of approximation algorithms in sublinear time.
It is applied to optimization problems for linear classifiers and minimum enclosing balls.
The algorithm employs a novel sampling technique along with a new multiplicative update procedure.
Hazan et al.~\cite{hazanbeating} exploited the approach to linear SVM (\textit{Support Vector Machine}).
Cotter et al.~\cite{cotter2012kernelized} continued to extend the method to kernelized SVM.
In~\cite{hazan2011optimal}, Hazan et al. applied the sublinear approximation approach for solving linear regression with penalties.
Garber and Hazan~\cite{garberapproximating} also developed it in SDP (\textit{Semidenfinite Programming}).
Peng et al.~\cite{peng2012sublinear} utilized the method in LR model with penalties and developed sequential sublinear algorithms for both $\ell_1$-penalty and $\ell_2$-penalty separately.

@@

\subsection{Logistic Regression Model}
In this paper, we are mainly concerned with the binary classification problem.  	
We define training dataset as $\XM=\{(\x_i, y_i): i=1, \ldots, n\}$, where $\x_i \in \RB^d$ are input samples and $y_i \in \{-1, 1\}$ are the corresponding labels.
For simplicity, we will use the notation $\X=[{\x}_{1}, {\x}_{2}, \ldots, {\x}_n]^{T}$ and $\y=(y_1, y_2, \ldots, y_n)^T$ to represent training dataset in the following.
Then to fit in the logistic regression model, the expected value of $y_i$ is given by
\[
P(y_i|\x_i)= \frac{1}{1+ \exp(- y_i(\x_i^T \w + b))} \triangleq g_i(y_i),
\]
where $\w=(w_1, \ldots, w_d)^T \in \RB^d$ is a regression vector and $b\in \RB$ is an offset term.

The learning process aims to optimize $\w$ and $b$.
To make the optimization problem more practical to solve, a common practice is to add penalties to LR model.
Typically, for $\ell_2$-penalty, we have to solve the following optimization problem:
\begin{equation} \label{eqn:1}
	\max_{\bw ,b} \; \Big\{F(\w, b| \XM) - \frac{\lambda}{2}  \|\bw\|_2^{2} \Big\}.
\end{equation}
For $\ell_1$-penalty, we are faced with the following optimization problem:
\begin{equation} \label{eqn:2}
	\max_{\bw ,b} \;  \big\{F(\w, b|\XM)- \gamma{\|\bw \|}_{1}\big\}.	
\end{equation}
Here, we have
\[
F(\w, b| \XM) = \sum_{i=1}^n \log g_i(y_i)
\]
in both (\ref{eqn:1}) and (\ref{eqn:2}).
To be brief, we omit the derivation here.

\subsection{Sequential Sublinear Algorithm}
We use the following notations to define sequential sublinear algorithm for penalized logistic regression.
Much of them are borrowed from~\cite{peng2012sublinear}.
$clip\lc \cdot \rc$ is a projection function defined as follows:
\[
clip\lc a,b \rc \triangleq \max \lc \min \lc a,b \rc ,-b\rc \,\,\,\,\,\, a,b \in \mathbb{R}.
\]

$\sgn \lc \cdot \rc$ is the sign function; namely,
\[
\sgn \lc x \rc =
\begin{cases}
	+1 & \text{ if } x>0  \\
	0  & \text{ if } x=0  \\
	-1 & \text{ if } x<0.
\end{cases}
\]

$g\lc \cdot \rc$ is the logistic function; namely,
\[
g \lc x \rc=\frac{1}{1+e^{-x}}
\]

In Algorithm~1, we give the sequential sublinear approximation procedure for logistic regression.	
	\begin{table} [ht]
	\begin{tabular}{l}
	\hline\noalign{\smallskip}
	\textbf{Algorithm 1} SLLR \\
	\noalign{\smallskip}
	\hline
	\noalign{\smallskip}
	1:	Input parameters: $\varepsilon, \nu~or~\gamma, X, Y$ \\
	2:	Initialize parameters: $T, \eta, {\mathbf{u}}_{0}, {\bw}_{1}, {\mathbf{\bq}}_{1}, {b}_{1}$\\
	3:	Iterations: $t=1 \sim T$ \\
	4:	\tspace ${\bp}_{t}\leftarrow{\bq}_{t}/{\|{\bq}_{t}\|}_{1}$ \\
	5:  \tspace Choose ${i}_{t}\leftarrow i$ with probability $\bp(i)$ \\
	6:	\tspace $coef\leftarrow{y}_{{i}_{t}}g\lc-{y}_{{i}_{t}}\lc {{\bw}_{t}}^{T}{\mathbf{x}}_{i_t}+{b}_{t} \rc\rc$ \\
	7:	\tspace ${\bu}_{t}\leftarrow {\bu}_{t-1}+\frac{coef}{\sqrt{2T}}{\mathbf{x}}_{{i}_{t}}$ \\
    8:  \tspace ${\bxi}_{t}\leftarrow \argmax_{\bxi\in \Lambda}\lc{{\bp}_{t}}^{T}\bxi\rc$, if input $\nu$ for $\ell_2$-penalty \\
    9:  \tspace $\bu_t\leftarrow$ soft-thresholding operations, if input $\gamma$ for $\ell_1$-penalty \\
	10:	\tspaces ${\bw}_{t}\leftarrow {\bu}_{t}/\max \left\{1,\|{\bu}_{t}\|_2 \right\}$ \\
    11: \tspaces ${b}_{t}\leftarrow \sgn\lc {{\bp}_{t}}^{T}\mathbf{y}\rc$ \\
	12:	\tspaces Choose ${j}_{t}\leftarrow j$ with probability ${{\bw}_{t}\lj}^{2}/{\|{\bw}_{t}\|_2}^{2} $ \\
	13:	\tspaces Iterations: $i=1 \sim n$ \\
	14:	\tspaces\tspace $\sigma \leftarrow \mathbf{x}_{i} \lc {j}_{t}\rc{\|{\bw}_{t}\|_2}^{2}/{\bw}_{t}\lc {j}_{t} \rc+{\bxi}_{t}\li+{y}_{i}{b}_{t}$ \\
	15:	\tspaces\tspace $\hat{\sigma} \leftarrow clip\lc \sigma,1/\eta \rc$ \\
	16:	\tspaces\tspace ${\bq}_{t+1}\li \leftarrow {\bp}_{t}\li\lc 1-\eta\hat{\sigma} + {\eta}^{2}{\hat{\sigma}}^{2} \rc$ \\
	17:	Output: $\bar{\bw}=\frac{1}{T}\sum_{t}{\bw}_{t},\bar{b}=\frac{1}{T}\sum_{t}{b}_{t}$ \\
	\hline
	\end{tabular} 	
	\label{alg:1}
	\end{table}
Each iteration of the \textbf{SLLR} algorithm have two phases: \textit{stochastic primal update} and \textit{stochastic dual update}.
In the pseudo-code of Algorithm 1, line 4 to line 11 is the primal part while line 12 to line 16 is the dual part.
We show the sublinear algorithm in a uniform way for both $\ell_2$-penalty and $\ell_1$-penalty concisely.
If we are dealing with $\ell_2$-penalty, we ignore line 9 and accomplish the computation in line 8 through a simple greedy algorithm.
Here, $\Lambda$ represents a Euclidean space with following conditions:
\[
\Lambda=\left\{ \bxi\in\mathbb{R}_{n} \, | \, \forall i, \, 0\leq{\xi}_{i}\leq 2, \, {\|\xi \|}_{1}\leq \nu n  \right
\}.
\]
If we are faced with $\ell_1$-penalty, we ignore line 8 and expand the procedure line 9 as following.
\begin{table} [ht]
\begin{tabular}{l}
\hline\noalign{\smallskip}
\textbf{Procedure} Line 9 in Algorithm 1\\
\noalign{\smallskip}
\hline
\noalign{\smallskip}
Iterations: $j=1 \sim d$ \\
\tspace \textbf{if} $\buprev_t\lj>0$ \textbf{and} $\bu_t\lj>0$ \\
\tspace\tspace $\bu_t\lj=\max \lc \bu_t\lj-\gamma ,0 \rc$ \\
\tspace \textbf{if} $\buprev_t\lj<0$ \textbf{and} $\bu_t\lj<0$ \\
\tspace\tspace $\bu_t\lj=\min \lc \bu_t\lj+\gamma ,0 \rc$ \\
$\buprev_{t+1}\leftarrow\bu_t$ \\
\hline
\end{tabular}
\end{table}

>>>>

% This is paper draft
% edited by Haoruo Peng
% Date: 2013.6.21
\documentclass[10pt, conference, compsocconf]{IEEEtran}
\usepackage{amsmath, amssymb}
\usepackage{color}
\usepackage{graphicx}
\usepackage{bbding}
\usepackage{epstopdf}

@@

\begin{document}

\title{Learning for Logistic Regression Model in Parallel}

\author{
\IEEEauthorblockN{Haoruo Peng}
\IEEEauthorblockA{HTC Research Lab\\Beijing, China\\penghaoruo@hotmail.com}
\and
\IEEEauthorblockN{Ding Liang}
\IEEEauthorblockA{HTC Research Lab\\Beijing, China\\Ding\_Liang@htc.com}
\and
\IEEEauthorblockN{Deli Zhao}
\IEEEauthorblockA{HTC Research Lab\\Beijing, China\\zhaodeli@gmail.com}
\and
\IEEEauthorblockN{Cyrus Choi}
\IEEEauthorblockA{HTC Research Lab\\Beijing, China\\Cyrus\_Choi@htc.com}
\and
\IEEEauthorblockN{Edward Y. Chang}
\IEEEauthorblockA{HTC Research Lab\\Beijing, China\\Edward\_Chang@htc.com}
}

\maketitle

\begin{abstract}
Logistic regression (LR) model has various applications in machine learning.
It is a supervised learning model for pattern classification and enjoys the benefits of statistical principles.
This paper addresses the issue of computational efficiency for solving LR in big data scenario.
We compare between different computing platforms and different parallel algorithms.
Our goal is to offer valuable suggestions on choosing computing platforms and learning algorithms according to different application situations for specific datasets and machine resources.
Three different parallel computing platforms along with three different types of parallel algorithms are carefully studied in this paper.
In such process, we present a novel parallel sublinear method based on its sequential version as another contribution.
%Hadoop, Mahout and Spark are three existing parallel computing platforms that can be implemented with large scale machine learning algorithms.
%Parallel gradient Descent and stochastic gradient descent are two state-of-art algorithms for solving LR.
To enhance practicality, we consider tradeoff between efficiency and precision.
We also provide fault tolerance for parallel distributed computations.
\end{abstract}

@@

\section{Introduction} \label{sec:int}
Logistic regression model~\cite{HastieBook:SL} plays a vital role in machine learning.
It is a widely used model in algorithms like PageRank~\cite{page1999pagerank} and anti-spam filtering~\cite{androutsopoulos2000evaluation}.
The model serves for classification problems, and is supported by a substantial body of statistical theories and algorithms.
A binary classification problem modeled by LR can be easily extended to a multi-class classification problem.
Also, due to the strong statistical foundation of the LR model, we can extend achieved results into a deeper analysis which may apply to other classification models. We will focus on the binary LR model in this paper.

In recent years, many modern datasets are increasingly larger and larger.
A most identifiable characteristic of these datasets is that data scale is very large and data dimensionality is very high.
Large data volume and high data dimensionality bring unavoidable computational challenges to machine learning problems.
For example, in social networks, we have datasets~\cite{lewis2008tastes} covering multiple dimensions like friendship relationship and share of common interests. They generally include millions of records over millions of attributes. More evidently, for the task of text categorization, we usually have to handle datasets with data volumes in billion scales and each data instance is characterized by thousands of feature dimensions~\cite{lewis2004rcv1}.
We tackle these challenges on both algorithm level and computing platform level via parallelization methods.
For many applications, in order to attain speedup in performance, we can employ not only task parallelism, but also data parallelism~\cite{subhlok1993exploiting}.
In this paper, we mainly discuss issues in task parallelism.

When developing parallel algorithms for LR, it is inevitable to come across the question of choosing which parallel computing platform to employ.
After a thorough survey, we choose three unique and best-known computing platforms to test: Hadoop~\cite{white2012hadoop}, Mahout~\cite{mahoutscalable} and Spark~\cite{zaharia2010spark}.
Hadoop employs HDFS~\cite{borthakur2008hdfs} and MapReduce~\cite{dean2008mapreduce}.
Mahout runs on Apache Hadoop and makes improvements in low level code.
Spark promotes efficiency of iterative algorithms and also supports HDFS.
We will compare design features of these three computing platforms separately in Section~\ref{sec:platform}.

In sequential optimization algorithms for LR model, a classical way is to resort to stochastic approximation methods.
Stochastic approximation methods, such as stochastic gradient descent~\cite{zhang2004solving} and stochastic dual averaging~\cite{xiao2010dual}, obtain optimal generalization guarantees with a small number of passes over data with runtime linear to the dataset size.
The stochastic gradient descent method can work in an online fashion, as we view sampled data points as data stream over algorithmic iteration.
It is hard to be changed to task-parallelism, but it is blazingly fast.
If we take a step back to the general gradient descent to solve LR model, we can find that it can be actually embarrassingly parallel.
We can take in all of the data at the same iteration and just compute the gradient in a MapReduce fashion.
We can further speed up the runtime by employing sublinear algorithms~\cite{peng2012sublinear} via the application of the stochastic approximation idea.
These methods access a single feature of feature vectors instead of entire feature vectors at each iteration.
We propose a parallel version of this algorithm and achieve the result of comparable accuracy while being faster convergent.

@@

\subsection{Scalability Solutions}
%
In big data scenarios, large datasets require us to develop machine learning algorithms towards a more efficient and more parallelized end.
Researchers have already done much work to cater for the needs of massive datasets.
Early work like Parallel Support Vector Machines (PSVM)~\cite{chang2011psvm} employed an approximation matrix decomposition method based on row selections to reduce memory usage.
It can then increase the number of parallel computational nodes to several hundreds.
Later, the work of Parallel Latent Dirichlet Allocation (PLDA)~\cite{wang2009plda} improves the computational efficiency of the LDA model by means of sampling methods.
The parallel algorithm in Hadoop is robust, for it has fault tolerance of machine failures, thus taking advantage of Hadoop features.
Recently, Dean \etal~\cite{le2011building} claimed that the advantage of parallelization is fully taken in deep learning algorithms.
It pushed the limit of parallel computational nodes to a hundred-million level, and in the meantime, achieved the best learning performance ever.
GraphLab~\cite{kyrola2012graphchi} is a recently developed tool for large-scale machine learning. It tackles efficiency and scalability problems of machine leaning algorithms by graph-theoretical models.

\subsection{Computing Platforms}\label{sec:platform}
Three computing platforms that we are working on have their own advantages to implement machine learning algorithms.
The Apache Hadoop~\cite{white2012hadoop} software library allows for distributed processing of large datasets across clusters of computers using simple programming models.
Hadoop utilizes MapReduce~\cite{dean2008mapreduce} as its computational paradigm.
MapReduce paradigm takes a set of key-value pairs as its input.
Then a user-defined \textit{map} function turns the input into some intermediate key-value pairs.
Another user-defined \textit{reduce} function merges values for the same intermediate pairs to produce output.
All of the operations are independent, thereby making MapReduce paradigm easy to be parallelized.
In addition, Hadoop provides a Distributed File System (HDFS).
Both MapReduce and HDFS are designed to handle node failures in an automatic way, making Hadoop support large clusters that are built on commodity hardware.
The goal of Apache Mahout~\cite{mahoutscalable} is to build scalable machine learning libraries. Their core algorithms for clustering, classification and collaborative filtering are implemented on top of Apache Hadoop using the MapReduce paradigm. The core libraries are highly optimized so that Apache Mahout also has good performance for non-distributed algorithms.
Spark~\cite{zaharia2010spark} is a cluster computing system that aims to make data analysis fast.
Spark supports in-memory cluster computing. A job can load data into memory and query it repeatedly by creating and caching resilient distributed datasets (RDDs).
Moreover, RDDs achieve fault tolerance through lineage: information about how RDDs are derived from other RDDs is stored in reliable storage, thus making RDDs easy to be rebuilt if a certain partition is lost.
Spark can run up to 100 times faster than Hadoop for iterative algorithms.

In Table~\ref{tab:systemcmp}, we compare different features between Hadoop and Spark computing platforms.
As Mahout is mainly built on Hadoop, it inherits most of Hadoop features.
For simplicity, we use short abbreviations in the table.
Spark supports two types of operations on RDDs: Actions and Transformations (As and Ts).
Actions include functions like \textit{count}, \textit{collect} and \textit{save}. They usually return a result from input RDDs.
Transformations include functions like \textit{map}, \textit{filter} and \textit{join}. They normally build new RDDs from other RDDs, which comply with the lineage rule.
In the table, we use "NFS" to represent "Normal File System" and "Fault Tolerance" here is specifically for node failure situations.
\begin{table}[h]
\centering
\caption{Platform Comparison: Hadoop vs. Spark}\label{tab:systemcmp}\vspace{-0.3cm}
\begin{tabular}{|c|c|c|}
\hline
           & Hadoop & Spark \\
\hline
Computational Paradigm  & MapReduce & As and Ts \\
\hline
File System Supported & HDFS & HDFS and NFS \\
\hline
Design Concept  & Key-value Pairs & RDD \\
\hline
Fault Tolerance Technique & Redundancy & Lineage \\
\hline
\end{tabular}\vspace{-0.5cm}
\end{table}

\subsection{Sublinear Methods}
Sublinear methods are recently proposed by many researchers.
Clarkson \etal~\cite{clarkson2010sublinear} first presented the solution of approximation algorithms in sublinear time.
It is applied to optimization problems for linear classifiers and minimum enclosing balls.
The algorithm employs a novel sampling technique along with a new multiplicative update procedure.
Hazan \etal~\cite{hazanbeating} exploited the approach to linear SVM.
Cotter \etal~\cite{cotter2012kernelized} extended the method to kernelized SVM.
In~\cite{hazan2011optimal}, Hazan \etal applied the sublinear approximation approach to solve linear regression with penalties.
Garber and Hazan~\cite{garberapproximating} also developed it in Semi-Definite Programming (SDP).
Peng \etal~\cite{peng2012sublinear} utilized the method in the LR model with penalties and developed sequential sublinear algorithms for both $\ell_1$-penalty and $\ell_2$-penalty.

@@

\subsection{Logistic Regression Model} \label{sec:def}
In this paper, we are mainly concerned with the binary classification problem.  	
We define the training dataset as $\XM=\{(\x_i, y_i): i=1, \ldots, n\}$, where $\x_i \in \RB^d$ are input samples and $y_i \in \{-1, 1\}$ are the corresponding labels.
For simplicity, we will use the notation $\X=[{\x}_{1}, {\x}_{2}, \ldots, {\x}_n]^{T}$ and $\y=(y_1, y_2, \ldots, y_n)^T$ to represent the training dataset in the following.
To fit in the logistic regression model, the expected value of $y_i$ is given by\vspace{-0.2cm}
\[
P(y_i|\x_i)= \frac{1}{1+ \exp(- y_i(\x_i^T \w + b))} \triangleq g_i(y_i),\vspace{-0.2cm}
\]
where $\w=(w_1, \ldots, w_d)^T \in \RB^d$ is a regression vector and $b\in \RB$ is an offset term.

The learning process aims to optimize $\w$ and $b$.
To make the optimization problem more practical to solve, a common practice is to add penalties to LR model.
Typically, for $\ell_2$-penalty, we solve the following optimization problem:\vspace{-0.2cm}
\begin{equation} \label{eqn:1}
	\max_{\bw ,b} \; \Big\{F(\w, b| \XM) - \frac{\lambda}{2}  \|\bw\|_2^{2} \Big\}.\vspace{-0.2cm}
\end{equation}
For $\ell_1$-penalty, we optimize the following criterion: \vspace{-0.2cm}
\begin{equation} \label{eqn:2}
	\max_{\bw ,b} \;  \big\{F(\w, b|\XM)- \gamma{\|\bw \|}_{1}\big\}.\vspace{-0.2cm}	
\end{equation}
Here, we define $F(\w, b| \XM) = \sum_{i=1}^n \log g_i(y_i)$ in both (\ref{eqn:1}) and (\ref{eqn:2}). To be brief, we omit the derivation here.

\subsection{Sequential Sublinear Algorithm}
We use the following notations to define sequential sublinear algorithm for penalized logistic regression. $clip\lc \cdot \rc$ is a projection function defined as $
clip\lc a,b \rc \triangleq \max \lc \min \lc a,b \rc ,-b\rc,\,a,b\in \mathbb{R}$. $\sgn \lc \cdot \rc$ is the sign function; namely, $\sgn \lc \cdot \rc \in \{-1,0,1\}$.
$g\lc \cdot \rc$ is the logistic function $g \lc x \rc=1/(1+e^{-x})$.
